---
title: "lab_1_gd"
author: "Gabriel De La Rosa"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)

```

Read in fisheries data

```{r}
us_landings <- read_csv(here("data", "noaa_fisheries.csv")) %>% 
  clean_names()
```

Let's clean this data frame up so it's tidy...

How do we get rid of the dollar sign?
tidyr::separate() then as.numeric()...
or readr::parse_number()

parse_number is awesome! use it within a mutate()
parse_number only displays 7 sig figs, but doesn't actually get rid of any digits


```{r}

landings_tidy <- us_landings %>% 
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_usd = (parse_number(dollars_usd,
                      na = c("NA"))))
```


Let's just get some salmon information...

How do we remove data that we don't want, aka the aggregate species data?
Use str_remove, and then specify by the string you want it to recognize. 
Careful, though, better have a good reason for ignoring that... jk this doesn't remove data, just the word...

We want data for just salmon. how do we filter by strings? use str_detect()
If you use it in the beginning of a filter function, it only keeps things with that string.

How do we separate species from salmon? we can use separate() 

```{r}

salmon_landings <- landings_tidy %>% 
  mutate(afs_clean = str_remove(afs_name, " aggregate")) %>% 
  filter(str_detect(afs_clean, patter = "salmon")) %>% 
  separate(afs_clean, into = c("group", "species"), sep = ",")

```


Find some grouped summary data:

let's find the annual total US landings and dollar value, summing across all states, for each TYPE of salmon using `group_by()` and `summarize()`

```{r}
salmon_summary <- salmon_landings %>% 
  group_by(year, species) %>% 
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_usd))
```


Make a graph...

```{r}

salmon_landings_graph <- 
  ggplot(salmon_summary, aes(x = year, y = tot_landings)) +
  geom_line(aes(color = species)) +
  theme_bw()
  
salmon_landings_graph

```

Now we can export our graph with ggsave()

```{r}

ggsave(plot = salmon_landings_graph, 
       here("figures", "us_salmon_gd.png"),
       height = 5, 
       width = 8)

```


Let's make a nice Kable table...


```{r}
salmon_first_5 <- salmon_summary %>% 
  head(5)
```

```{r}

kable(salmon_first_5) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)

```

